Liveness-Based Garbage Collection for Lazy Languages

### 2016-02-13 paper/paper.tex
in 12 pages...
fixed for compliance with ISMM
spell checks

\documentclass[9pt,preprint]{sigplanconf}
\usepackage{etex}
\usepackage{float}
\usepackage{mathrsfs}
%%\usepackage{lineno}
\usepackage{booktabs}
\usepackage{stmaryrd}
\usepackage{boxedminipage}
\usepackage[T1]{fontenc}
\usepackage{epsfig}
\usepackage{multirow}
\usepackage{url}
\usepackage[normalem]{ulem}
\usepackage{schemepgm}
\usepackage{graphicx}
\usepackage{graphpap}
\usepackage{tabularx}
\usepackage{amssymb}
\usepackage{amsmath}
\usepackage{pstricks}
\usepackage{pst-text}
\usepackage{pst-node}
\usepackage{pst-tree}
\usepackage{pst-rel-points}
\usepackage{bcprules}
\usepackage{subfigure}
\usepackage[boxed]{algorithm2e}
\usepackage{makecell}
\usepackage{balance}

@@

\definecolor{Myblue}{rgb}{.2,0,1}
\definecolor{Myred}{rgb}{1,.2,.2}
\newcommand{\mycomment}[1]{{\color{Myblue}{#1}}}
\newcommand{\added}[1]{{\color{Myblue}{#1}}}
%\newcommand{\added}[1]{#1}
\newcommand{\todelete}[1]{}%{\color{Myred}{#1}}}

%\newcommand{\warning}[1]{{\color{Myred}{#1}}}
\newcommand{\warning}[1]{}
%\newcommand{\blankout}[1]{}
\newcommand{\oxlb}{\ensuremath{[\!|}}
\newcommand{\oxrb}{\ensuremath{|\!]}}
\newcommand{\cred}[1]{\psframebox[linestyle=none, fillcolor=lightgray,fillstyle=solid,framesep=0.5pt]{#1}}
%\newcommand{\deltacall}[3]{\delta_{#1}({#2},{#3})}
\newcommand{\scmin} {\mbox{\sf\em in}}
\newcommand{\scmnull}{\mbox{\sf\em null?}}
\newcommand{\scmpair}{\mbox{\sf\em pair?}}
\newcommand{\scmprim}{\ensuremath{\mathsf{+}}}
\newcommand{\sh}[1]{{\colorbox{gray!20}{\framebox{$#1$}}}}
\def\myvec{\mathaccent"017E } 
\newcommand{\stk}{{S}}      
\newcommand{\ID}{\mbox{$\mathbf{ id}$}}
\newcommand{\bang}{\mbox{\sc bang}}
\newtheorem{theorem}{Theorem}[section]
\newtheorem{proposition}[theorem]{Proposition}
\newtheorem{definition}[theorem]{Definition}
\newtheorem{lemma}[theorem]{Lemma}
\begin{document}
%% \linenumbers 
% --- Author Metadata here ---
%\conferenceinfo{WOODSTOCK}{'97 El Paso, Texas USA}
%\CopyrightYear{2007} % Allows default copyright year (20XX) to be
%over-ridden - IF NEED BE.
%\crdata{0-12345-67-8/90/01}  % Allows default copyright data
%(0-89791-88-6/97/05) to be over-ridden - IF NEED BE. 
% --- End of Author Metadata ---

@@

\begin{table*}[t!]
\caption{Statistics for liveness analysis and garbage collection}
\label{tab:exp-results}
\centering
\include{data-liveness-analysis}
\vskip -5mm
\include{autotable}
\vskip -5mm
\end{table*}

@@

There have been  several attempts to use liveness  analysis to improve
GC for imperative  languages.~\cite{khedker07heap} presents a liveness
analysis and uses  the results for inserting  nullifying statements in
Java  programs. In~\cite{ran.shaham-sas03},  temporal properties  like
liveness  are checked  against an  automaton modelling  heap accesses.
Both these approaches are intraprocedural in scope.

In the space of functional languages, there are: rewriting methods
such as
deforestation~\cite{wadler88deforest,gill93ashort,chitil99deforest},
sharing analysis based reallocation~\cite{jones89compile}, region
based analysis~\cite{tofte98region}, and insertion of compile-time
nullifying statements~\cite{inoue88analysis, lee05static}.  All compile-time
marking approaches rely on an efficient and precise alias analysis and
cannot provide significant improvement in its absence. The only work
in the space of lazy languages seems to be~\cite{Hamilton}
which touches upon only basic techniques of compile-time garbage
marking, explicit deallocation and desctructive allocation. An
interesting approach suggested in~\cite{HofmannJ03} is to annotate the
heap usage of first-order programs through linear types.  The
annotations are then used to serve memory requests through
re-allocation.  However, this requires the user to write programs in a
specific way.
%
Safe-for-space~\cite{appel.cps}              approaches~\cite{Clinger,
  shao00efficient}  reduce the  amount of  heap used  by a  program by
allocating   closures    in   registers   and   through    tail   call
optimizations. However, these approaches take care of only part of the
problem addressed  by our  analysis as the  program can  still contain
unused objects and closures that are reachable.
%

>>>>

\documentclass[9pt,preprint,letter,nonatbib]{sigplanconf}
\usepackage{etex}
\usepackage{float}
\usepackage{mathrsfs}
%%\usepackage{lineno}
\usepackage{booktabs}
\usepackage{stmaryrd}
\usepackage{boxedminipage}
\usepackage[T1]{fontenc}
\usepackage{epsfig}
\usepackage{multirow}
\usepackage{url}
\usepackage[normalem]{ulem}
\usepackage{schemepgm}
\usepackage{graphicx}
\usepackage{graphpap}
\usepackage{tabularx}
\usepackage{amssymb}
\usepackage{amsmath}
\usepackage{pstricks}
\usepackage{pst-text}
\usepackage{pst-node}
\usepackage{pst-tree}
\usepackage{pst-rel-points}
\usepackage{bcprules}
\usepackage{subfigure}
\usepackage[boxed]{algorithm2e}
\usepackage{makecell}
\usepackage{balance}

@@

\definecolor{Myblue}{rgb}{.2,0,1}
\definecolor{Myred}{rgb}{1,.2,.2}

\newcommand{\mycomment}[1]{{\color{Myblue}{#1}}}
%\newcommand{\added}[1]{{\color{Myblue}{#1}}}
\newcommand{\added}[1]{#1}
\newcommand{\todelete}[1]{}%{\color{Myred}{#1}}}
%\newcommand{\warning}[1]{{\color{Myred}{#1}}}
\newcommand{\warning}[1]{}
%\newcommand{\blankout}[1]{}

\newcommand{\oxlb}{\ensuremath{[\!|}}
\newcommand{\oxrb}{\ensuremath{|\!]}}
\newcommand{\cred}[1]{\psframebox[linestyle=none, fillcolor=lightgray,fillstyle=solid,framesep=0.5pt]{#1}}
%\newcommand{\deltacall}[3]{\delta_{#1}({#2},{#3})}
\newcommand{\scmin} {\mbox{\sf\em in}}
\newcommand{\scmnull}{\mbox{\sf\em null?}}
\newcommand{\scmpair}{\mbox{\sf\em pair?}}
\newcommand{\scmprim}{\ensuremath{\mathsf{+}}}
\newcommand{\sh}[1]{{\colorbox{gray!20}{\framebox{$#1$}}}}
\def\myvec{\mathaccent"017E } 
\newcommand{\stk}{{S}}      
\newcommand{\ID}{\mbox{$\mathbf{ id}$}}
\newcommand{\bang}{\mbox{\sc bang}}
\newtheorem{theorem}{Theorem}[section]
\newtheorem{proposition}[theorem]{Proposition}
\newtheorem{definition}[theorem]{Definition}
\newtheorem{lemma}[theorem]{Lemma}
\begin{document}

\special{papersize=8.5in,11in}
\setlength{\pdfpageheight}{\paperheight}
\setlength{\pdfpagewidth}{\paperwidth}

%% \linenumbers 
% --- Author Metadata here ---
%\conferenceinfo{WOODSTOCK}{'97 El Paso, Texas USA}
%\CopyrightYear{2007} % Allows default copyright year (20XX) to be
%over-ridden - IF NEED BE.
%\crdata{0-12345-67-8/90/01}  % Allows default copyright data
%(0-89791-88-6/97/05) to be over-ridden - IF NEED BE. 
% --- End of Author Metadata ---

@@

\begin{table*}[t!]
\caption{Statistics for liveness analysis and garbage collection}
\label{tab:exp-results}
\centering
\include{data-liveness-analysis}
\vskip -5mm
\include{autotable}
\vskip -7mm
\end{table*}

@@

There have been  several attempts to use liveness  analysis to improve
GC for imperative  languages.~\cite{khedker07heap} presents a liveness
analysis and uses  the results for inserting  nullifying statements in
Java  programs. In~\cite{ran.shaham-sas03},  temporal properties  like
liveness  are checked  against an  automaton modeling  heap accesses.
Both these approaches are intraprocedural in scope.

In the space of functional languages, there are: rewriting methods
such as
deforestation~\cite{wadler88deforest,gill93ashort,chitil99deforest},
sharing analysis based reallocation~\cite{jones89compile}, region
based analysis~\cite{tofte98region}, and insertion of compile-time
nullifying statements~\cite{inoue88analysis, lee05static}.  All compile-time
marking approaches rely on an efficient and precise alias analysis and
cannot provide significant improvement in its absence. The only work
in the space of lazy languages seems to be~\cite{Hamilton}
which touches upon only basic techniques of compile-time garbage
marking, explicit deallocation and destructive allocation. An
interesting approach suggested in~\cite{HofmannJ03} is to annotate the
heap usage of first-order programs through linear types.  The
annotations are then used to serve memory requests through
re-allocation.  However, this requires the user to write programs in a
specific way.
%
Safe-for-space~\cite{appel.cps}              approaches~\cite{Clinger,
  shao00efficient}  reduce the  amount of  heap used  by a  program by
allocating   closures    in   registers   and   through    tail   call
optimizations. However, these approaches take care of only part of the
problem addressed  by our  analysis as the  program can  still contain
unused objects and closures that are reachable.
%
